Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-24 Multiresolution_framework/ltx/manuscript.tex
small edit to IDE neural field model section to generalise it

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. On finer scales... where any measurable spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location \parham{with} the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field.

This multi-resolution structure has been described by numerous methods \cite{Wilson1973}. This idea of multi-resolution can be thought of a generalisation of modelling the layers of the cortex separately, where the connectivity kernel, or neural foot-print, has varying widths.

Typically, the experimentalist only has access the net field generated by the multi-resolution architecture. Follow this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \cite{Breakspear2005} in using a wavelet decomposition.  
\begin{figure}[t]
	\centering
		\includegraphics[scale=1]{./Graph/MultiResNeuralField.pdf}
	\caption{{\bf Multi-layer neural field model}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

\section{IDE Neural Field Model}

\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ (width $\sigma_m$) & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$, $\xi$ & inverse synaptic time constant \& time constant parameter & s$^{-1}$, n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$\textrm{D}$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal K_{t} $ & filter gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}
\doublespacing

The stochastic IDE form of the Amari neural field  formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the membrane time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Figure~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=.6]{./Graph/ConnectivityKernel.pdf}
\includegraphics[scale=.6]{./Graph/Synaptickernel.pdf}
\includegraphics[scale=.6]{./Graph/Sigmoid.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{a}) The mexican hat connectivity kernel. (\textbf{b}) First order synaptic kernel. (\textbf{c}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}

>>>>

\section{Multi-Resolution Cortical Architecture}
The human neocortex has an inherent multi-resolution architecture. The spatiotemporal neural field can be written as 
\begin{equation}
	v\left(\mathbf{r},t\right) = \sum_{j} v_j\left(\mathbf{r},t\right),
\end{equation}
where $v\left(\mathbf{r},t\right)$ is a voltage field that may be measured with a multi-electrode array or voltage sensitive dye, $\mathbf{r}\in \mathbb{R}^n$ (where $n\le3$) describes the spatial location within the field, $t$ denotes time, and $j$ indexes the different levels of resolution in the field.

This multi-resolution structure has been described by numerous methods. For example, this idea of multi-resolution can be thought of a generalisation of modelling the layers of the cortex separately, where the connectivity kernel, or neural foot-print, has varying widths \cite{Wilson1973}.

Typically, the experimentalist only has access the net field generated by the multi-resolution architecture. Follow this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \cite{Breakspear2005} in using a wavelet decomposition. This paper is directed at developing a framework for patient-specific multi-resolution neural field models. 
\begin{figure}[t]
	\centering
		\includegraphics[scale=1]{./Graph/MultiResNeuralField.pdf}
	\caption{{\bf Multi-layer neural field model}. The inset figures show the shape of the
connectivity kernel components at each spatial resolution.}
	\label{fig:MultiLayerFieldModel}
\end{figure}

\section{IDE Neural Field Model}
\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ (width $\sigma_m$) & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$, $\xi$ & inverse synaptic time constant \& time constant parameter & s$^{-1}$, n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$\textrm{D}$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal K_{t} $ & filter gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}
\doublespacing

The neural field model in the most general form is described by the equation
\begin{equation}
	\label{FullDoubleIntModel} \underbrace{v\left(\mathbf{r},t\right)}_{field} =
	\int_{-\infty}^t 
	\underbrace{h\left(t - t'\right)}_{\text{synapse}} \int_\Omega
	\underbrace{w\left(\mathbf{r},\mathbf{r}'\right)}_{\text{connectivity}} 
	\underbrace{f\left( v\left( \mathbf{r}',t' \right)\right)}_{\text{firing}}
	\, \mathrm{d}\mathbf{r}'\mathrm{d}t',
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted by $v_t\left(\mathbf r\right)$, $h(t)$ describes the post-synaptic response kernel, $w(\mathbf{r})$ describes the connectivity kernel and $f(v(\mathbf{r},t))$ describes the relationship between the mean membrane potential and the mean firing rate. There are various forms of $h(t)$ that have been proposed in the literature, with different levels of complexity. These are shown in Figure~\ref{fig:Modelcomponents}. Similarly, both the connectivity kernel and the activity function (firing rate) can take various forms. These are also shown in Figure~\ref{fig:Modelcomponents}.

The stochastic IDE form of the Amari neural field formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, \mathrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where in this first-order (in time) form of the model the membrane dynamics are included through the parameter $\xi=1-T_s/\tau$, where $\tau$ is the membrane time constant and $T_s$ is the sampling time. 
% The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the connectivity kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977,Atay2005}. The components of the model are shown in Figure~\ref{fig:Modelcomponents}
\begin{figure}[!t]
\centering
\includegraphics[scale=.6]{./Graph/ConnectivityKernel.pdf}
\includegraphics[scale=.6]{./Graph/Synaptickernel.pdf}
\includegraphics[scale=.6]{./Graph/Sigmoid.pdf}
\caption{ {\bf The IDE neural field model components}. (\textbf{a}) The mexican hat connectivity kernel. (\textbf{b}) First order synaptic kernel. (\textbf{c}) A representative of the activation function.}
\label{fig:Modelcomponents}
\end{figure}
